---
title: "Module 3"
output: html_document
date: "2023-11-24"
---


#Loading required packages----
```{r}
library(base)

library(tidyverse)

library(dplyr)
```

```{r}
property_assessment<- read.csv("fy2023-property-assessment-data.csv")
```

```{r}
library(janitor)

property_assessment<- clean_names(property_assessment)
```

```{r}
updated_property_assessment<-  property_assessment[, c("city", "lu", "lu_desc", "bldg_type", "res_floor", "res_units", "com_units", "rc_units", "gross_area", "living_area", "land_value", "bldg_value", "total_value", "gross_tax", "yr_built", "bed_rms", "full_bth", "hlf_bth", "kitchens", "bdrm_cond", "heat_type", "ac_type", "num_parking", "prop_view")]


```

```{r}
int_columns <- sapply(updated_property_assessment, is.integer)

updated_property_assessment[, int_columns] <- lapply(updated_property_assessment[, int_columns], as.numeric)

str(updated_property_assessment)

```

```{r}
clean_num_columns <- function(x) {
  # Remove outliers based on IQR
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr_value <- IQR(x, na.rm = TRUE)
  
  lower_bound <- q1 - 1.5 * iqr_value
  upper_bound <- q3 + 1.5 * iqr_value
  
  x_outliers_removed <- ifelse(x < lower_bound | x > upper_bound, NA, x)
  
  # Impute missing values with the median
  x_clean_data <- ifelse(is.na(x_outliers_removed), median(x, na.rm = TRUE), x_outliers_removed)
  
  return(x_clean_data)
}

#Clean all numeric columns in the dataset
updated_property_assessment <- updated_property_assessment %>%
  mutate(across(where(is.numeric), clean_num_columns))

# Remove rows with any NA values after cleaning
updated_property_assessment <- na.omit(updated_property_assessment)

updated_property_assessment <- updated_property_assessment %>%
  mutate(city = ifelse(city == "" | is.na(city), "Unknown", city))

```


#Calculating t tests----

#Considering variable Gross_tax----
```{r}

mean_gross_tax<- mean(updated_property_assessment$gross_tax)

print(mean_gross_tax)
```
Null Hypothesis: Mean Gross Tax is equal to 6187.053
Alternative Hypothesis: Mean Gross Tax is not equal to 6187.053

```{r}

gross_tax<- updated_property_assessment$gross_tax


set.seed(123)

sample_gross_tax<- sample(gross_tax, 200)

mean_sample_gross_tax<- mean(sample_gross_tax)

# One-sample t-test
t_test_gross_tax <- t.test(sample_gross_tax, mu = mean_gross_tax)

# Calculation of t(critical)
confidence_level <- 0.95  
degrees_of_freedom <- 199

# Calculate the critical t-value
t_critical <- qt((1 + confidence_level) / 2, df = degrees_of_freedom)

# Print the result

# Print the results
cat("Null Hypothesis (H0): The mean of the sample is equal to the population mean.\n")
cat("Alternative Hypothesis (H1): The mean of the sample is not equal to the population mean.\n\n")


cat("T-test Results:\n")
print(t_test_gross_tax)
cat("t(critical) for 95% confidence interval and 99 degree of freedom = ",t_critical)


```
Since the value of t is less than the value of t critical, we do not have enough evidence to reject our Null Hypothesis.
95 confidence interval is 5378.219 6516.200
True population mean lies within 6187.053


#Considering variable Building Value----
```{r}

mean_bldg_value<- mean(updated_property_assessment$bldg_value)
print(mean_bldg_value)

```

```{r}
bldg_value<- updated_property_assessment$bldg_value


set.seed(100)

sample_bldg_value<- sample(bldg_value, 200)

mean_sample_bldg_value<- mean(sample_bldg_value)

# One-sample t-test
t_test_bldg_value <- t.test(sample_bldg_value, mu = mean_bldg_value)

# Calculation of t(critical)
confidence_level <- 0.95  
degrees_of_freedom <- 199


# Calculate the critical t-value
t_critical_bv <- qt((1 + confidence_level) / 2, df = degrees_of_freedom)

cat("Null Hypothesis (H0): The mean of the sample is equal to the population mean of bldg_value.\n")
cat("Alternative Hypothesis (H1): The mean of the sample is not equal to the population mean of bldg_value.\n\n")


cat("T-test Results:\n")
print(t_test_bldg_value)

cat("t(critical) for 95% confidence interval and 99 degree of freedom = ",t_critical)
```
Given the p-value of 0.4692 we do not have much evidence to reject the Null Hypothesis.The t-test suggests that the sample mean is not significantly different from the assumed population mean of bldg_value. The confidence interval also supports this conclusion, as it includes the assumed population mean.
95 percent confidence interval is 447914.6 527171.6
True population mean lies within 472968.9
